### 2015-08-14 tex/dtree.tex
Updates from ShareLaTeX

\begin{figure}[t]
	\small
	~\hrule~

	{\bf Step 1: Distinguish between clusters using  decision trees}

	Use an entropy-based
	decision tree (DT) learner to learn what attributes select for a given data. To limit tree size:
	\bi
	\item Only use the top $\alpha=33$\%  of the features, as determined by their information gain~\cite{Irani1993}. 
	\item Only build the trees down to  max depth of $\beta=10$.
	\item Only build subtrees if it contains at least $N^{\gamma=0.5}$ examples (where $N$ is the size of the training set).
	\ei
	{\em Score}  DT  leaf nodes  via the mean score of its majority cluster.

	~\hrule~

	{\bf Step 2: Generating contrast sets from DT branches}
	\begin{itemize}
		\item Find the {\em current } cluster: take each test instance, run it down to a leaf in the DT tree.  
		\item Find the {\em desired} cluster: 
		\bi
		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
		\item Identify {\em sibling} clusters; i.e. leaf clusters that can be reached from level $lvl$ that are not {\em current }
		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\epsilon=0.5$ times the mean score of {\em current}. If none found, then repeat for $lvl += 1$
		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
		\ei
		\item Find the {\em delta}; i.e. the set difference between  conditions in the DT branch to {\em desired} and {\em current}. To find that delta:
		\bi
		\item
		For discrete attributes,  return the value from {\em desired}. 
		\item
		For  numerics, return the numeric difference. 
		\item
		For numerics  into ranges, return a random number selected from the low and high boundaries of the that range.
		\ei
		\ei
		~\hrule~
		\caption{CROSSTREES. Controlled by the parameters
			$\{\alpha, \beta, \gamma, \delta, \epsilon\}$ (set via engineering judgement).}
		\label{fig:xtrees_bare}
	\end{figure}