### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

\documentclass{sig-alternate}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{cite}
\usepackage{framed,graphicx,xcolor}

\usepackage{stfloats}

\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
%\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\usepackage{program}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
%\usepackage[table]{xcolor}
\definecolor{darkgreen}{rgb}{0,0.3,0}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\usepackage{colortbl}
\usepackage{picture}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{listings}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Comments}{rgb}{0.5,0.5,0.5}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{1,1,1}

 \definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
\newcommand{\G}{\cellcolor{green}}
\newcommand{\Y}{\cellcolor{yellow}}

\newcommand{\quart}[4]{\begin{picture}(100,4)%1
{\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}

\definecolor{MyDarkBlue}{rgb}{0,0.08,0.45} 
\newenvironment{changed}{\par\color{MyDarkBlue}}{\par}
%\newenvironment{changed}{\par}{\par}

\newcommand{\ADD}[1]{\textcolor{MyDarkBlue}{{\bf #1}}}
\usepackage{times}
\pagenumbering{arabic} 
\begin{document}

\definecolor{shadecolor}{gray}{0.9}
\conferenceinfo{ISCE}{'16 Austin, Texas}

\title{How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)}
\numberofauthors{3} 
\author{
\alignauthor 
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{Com. Sci., NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Naveen   Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Washington D.C.}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}
}
\maketitle
\begin{abstract}
 As business users demand more insightful
 analytics, data scientists need to change
 their tools. Instead of just predicting 
 some result, they also need tools that generate ``plans'';
 i.e. specific suggestions on  what to change  in order to
 improve on the predicted values.
This paper defines an evaluation strategy for evaluating such  planners and applies it to four different planning methods.
It is shown that our  strategy  can (a) recognize   inferior planning methods and  (b) select useful planning methods.
Using the best method found by this evaluation strategy,  we can
reduce
 the expected values of defect counts and  runtimes of software projects to    
 45\%,61.5\%  (median) and 22\%,6\% (best case) of the initial values, respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} approach called XTREES for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREES
to reduce  the expected value of:
\bi
\item  Defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10}.
\item Runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.
\ei 
The approach discussed here is novel since most data miners used in software analytics
comment on ``what is'' and not ``what to do''. For example, a clustering
algorithm might find clumps of similar examples, but does not comment on the minimal
change(s) required to move from clump to clump.

The contributions of this paper are two-fold:
\be
\item
The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
can determine that some rows are preferred and some are to be avoided).  
\item An evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. With this strategy, other researchers could assess improvements or alternatives to the XTREE planner. 
\ei
The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREES. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

\section{Preliminaries}\label{sect:prelim}
\subsection{From Prediction to Planning}

This paper is about the next step {\em after} prediction. Suppose
a business user is presented a prediction and they do not like what they see; e.g. the runtimes are too long of the number of defects is too high. This user may
then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''

Before exploring automatic methods to answer the planning question, we first comment
on two manual methods.

One way to propose changes to a project would be to   ask some smart experienced
person for their opinion on how to (e.g.) reduce defects and/or decrease runtimes. Sometimes
such advice 
is an effective strategy and sometimes it is not.
According to Passos et al.~\cite{passos11},  developers
may  assume that the lessons they learn from a few past
projects are general to 
all their future projects. They comment ``past experiences were taken into account without 
much consideration for their context~\cite{passos11}.  
 J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that 
  supposed software engineering    ``gurus'' rarely use lessons
  from past projects to improve their future reasoning and that such poor
  past advice can be detrimental to new projects.~\cite{jorgensen09}.
  Accordingly, we   propose a ``trust, but verify'' approach.
  After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
  if there are any better options
  (just as a sanity check).
  The rest of this paper discusses some methods to build automatic oracles 
  to implement that   sanity check.

Another way  to find   changes to a project
might be to rely
on the peer review processes used by the 
SE research community. This approach would propose changes to software
projects that concur with internationally accept best practices. 
There are two
problems with that approach. Firstly, given the rapid pace of change in software
engineering, we may be asking questions for which there is no currently accept
``best practice''. For example, only very recently has there been any work
of predicting software runtimes based on choices within Makefiles. While the work
of Siegmund et al.~\cite{sven12} is certainly state of the art, it does
cannot yet represent ``internationally accept best practices'' since that work is so
recent.

Secondly, given the diversity of SE products and practices
and personnel, it may well be that the current project being discussed is 
substantively different to prior work. 
Numerous recent {\em local learning} results compare (1) single models
learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
A repeated result in those studies in that the local models generated the better effort
and defect predictions (better median results,
lower variance in the predictions). This paper offers yet another locality result:
\bi
\item
One standard rule in the literature
is that it is useful to implement modules such that they are internally cohesive (use
much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
\item
While that may be true in general, for particular classes other changes may be more important
(later in this paper, we show one set of results were that is indeed the case).
\ei
In summary, 
it is useful to have automatic methods to recommend changes. Such
methods can fill in for human gurus (if such gurus are absent) or 
to offer a second opinion.
Also, prior to making automatic recommendations, it is wise to first stratify the data
(clump it into related examples) then generate advice specific to each clump.
Accordingly, the rest of this paper defines and evaluates
automatic methods to find plans from
  $N$ examples divided  into  many clumps.

\input{tex/jur_char.tex}

\subsection{Trusting the Changes}\label{sect:trust}
   XTREES is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never seen before. Therefore, it must be asked:
\bi
\item
{\em Can we trust the predictions made on such new examples?}
\ei
Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
any data miner, there are two ``clouds'' of data:
\bi
\item One cloud is the space of training examples.
\item Another cloud is the space of test examples.
\ei
For a visualization of these clouds, see \fig{howxy}.

We should mistrust the predictions made by a  model   if it is being applied to examples far outside of its region of competency;
i.e. to test examples that are too far removed from the training data.   
To test for ``too far'', we can run a data mining experiment to test how well
a model learned from the training data applies to the test data. Such experiments return some performance value.

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
Given this, one thing  can be asserted about predictions on changed examples:
\bi
\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

This pattern of \fig{howxy} (where the displaced test cases are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
paper. Hence,  we can assert that
predictors learned from these training data have some authority in the regions
used to generate the plans\footnote{If the reader is concerned that this 2d projection
losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
have confirmed that building defect predictors from this 2d space are just as effective
as predictors learned using all the defect attributes of \fig{ck}.}.

\begin{figure}
\includegraphics[width=2in]{_figs/twod.eps} 
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}
%XXX end

 \begin{figure}[!t]
 \small
 \begin{center}
 \begin{tabular}{r|rr}
 data set & cases & \% defective\\\hline
  ant &947& 22\\
  camel& 1819& 19\\
 jedit& 1257& 2\\
 ivy &352& 11\\
 log4j& 244 &92\\
 lucebe &442 &59\\
 poi& 936 &64\\
 synapse &379 &34\\
 velocity& 410& 34\\
 xalan& 2411& 99\\
 xerces &1055& 74
 \end{tabular}
 \end{center}
 \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
 \end{figure}

 \begin{figure}[!t]
\scriptsize
\begin{tabular}{llllll}
  \hline
  \rowcolor{lightgray}
Project & Domain & Lang. & LOC & Features & Config\\\hline
Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\
Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\
Apache & Web Server & C & 230,277 & 9 & 192\\
SQLite & Database & C & 312,625 & 39 & 3,932,160\\
LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
\end{tabular}

\caption{Siegmund data.
For SQLite, the data  contains 4,553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}\label{fig:cpm}
\end{figure}

\begin{figure}[!t]
\includegraphics[width=1\linewidth]{_figs/BDBC.eps}
\caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
\end{figure}

That said, the above comes with two important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the next section. 
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see the Validity section, later in this paper.
\ei

  \section{Test Data}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

   The Jurecko data records number of known defects for each class, where the classes are described in terms of
  nearly two dozen metrics such as number of children, lines of code, etc. For details on the Jurecko data, see  \fig{ck} and \fig{jd}. 
 For the most part, the methods of this paper treat
  the Jurecko as a discrete class data set, where {\em defects} are true if the raw defect count is greater than zero.
  The one exception will be the `best-in-cluster'' method that reflects on the numeric value of the raw defect count (see \tion{BIC}).

  The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To make that data, Siegmund et al. perturbed
  the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
  Berkeley database (one written in ``C'' and one in Java). 
   \fig{bdbc} shows an example of a feature model defining valid combinations of settings to on the
   the Siegmund et al. datasets. These feature models were used by Siegmund et al. to ensure all their perturbations are value
   (we will use the same models to cull invalid plans).
  Given those valid perturbations, the systems were then compiled and 
  Siegmund et al. recorded how long each perturbation took to run a test suite.

\input{tex/prediction_stats.tex}

  Our evaluation strategy (discussed below) divides this data into a training a test set.
  From the train set we apply a data miner (to learn a quality predictor) and
  various planning methods (to learn different plans). Next, we try applying
  each of those plans to the  test set and ask the quality predictor to assess the changed examples.
  Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
  in the changed examples.

  As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
  For the Siegmind data, this criteria was   relatively easy to achieve.
  The data in those data sets have a continuous class (runtime of the compiled system)
  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
  x264. 
  As a preliminary study, we split that data   into equal sized train:test groups
  and trained a Random Forest
  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
That is, we can be very confident that the predictors from the Siegmund data can assess
our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

 It proved to be  more complicated to commission the Jureczko data sets for this study.
 For that data, we found that the
 quality predictors built from this data are far from perfect;
However, for some data sets, the  predictors could
be salvaged using the techniques discussed in this section.

 \fig{j} shows our preliminary studies with the Jureczko   data.
Given access to $V$ released
versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
shown in \fig{j}, this means that we are training on hundreds to thousands
of classes and testing on smaller test suites).
Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately
defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish
good from bad (since there are so many bad examples).

The  Jureczko data uses non-numeric discrete classes (``defective'' or ``not'').
For such data, quality predictor   is be measured using:
\bi
\item Probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
the test data detected
by the {\em predictor}.
\item Probability of false alarm (a.k.a. ``pf''): the percent of non-fault
classes that are {\em predicted} to be defective.
\ei 
As a preliminary study, we split the Jureczko  data   into equal sized train:test groups.
Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
built from the training data, then applied to the test data.
The ``untuned'' columns of \fig{j} shows those results.
If we define ``good'' to mean $\mathit{pd}>60 \wedge \mathit{pf} < 40$\%,
then only two of our data sets ({\em ivy,ant}) are ``good'' enough for this study.
Note that, as might have been expected, none of the \colorbox{lavenderpink}{three bottom rows} of \fig{j} were ``good''.

Fortunately,
the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets. Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples. Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
for tree generation, etc. The rows marked with $\star$ in \fig{j} show data sets whose performance was improved remarkably by these techniques. For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%. However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.

In summary, while we cannot trust predictors from some of our defect data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
Accordingly, when this study explores the Jureczko data, we will use these five data sets.

(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
parameter tunings, those techniques were applied to the training data and {\em not}
the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)

	\begin{figure}[!t]
	\begin{shaded}
  ~\hrule~

 {\bf Figure 5.A: Measuring Variability}

	For  continuous and discrete values,
the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
for  $n$ discrete values  at frequency
$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.

  ~\hrule~

	{\bf Figure 5.B: Measuring distance}

	We use
		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  \Delta(X_i,Y_i)^2}/\sqrt{F}$
			(the denominator normalizes the output to $0 \le d \le 1$).
			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
			Otherwise, replace any  missing items with values that maximizes the following.
			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
			returns 
			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
			same,different (respectively).

 ~\hrule~

	{\bf Figure 5.C: Top-down Clustering with WHERE}

				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
		Using this measure, WHERE runs as follows:
		\begin{enumerate}[leftmargin=3mm]
			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(which requires only $O(2N)$ comparisons).
			\item Project each case $Z$
			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
			\item Split the data at the median $x$ value of all cases.
			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse   from step1.
		\end{enumerate}
 In terms of related work,
	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
	  at each recursive level to find the PCA principle component. Our method, on the other hand,
	  performs the same task with only $O(2N)$ distance calculations 	using the 
	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.

	   ~\hrule~

	{\bf Figure 5.D: Top-down division with Decision Trees}

	Find a split in the values of  independent features that most reduces the variability
  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.

	 ~\hrule~
	  \end{shaded}
		\caption{Some algorithms used in this paper.}\label{fig:where}
		\end{figure}

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
  Note that all our methods have the two properties proposed in \tion{prelim}:
  \bi
  \item They find different plans in different parts of the data;
  \item They  are {\em trust-increasing} in the manner described in \tion{trust};
  i.e. they change changing examples such that they move {\em closer} to the training data.
  \ei
The following descriptions of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

\subsection{Method1 = CD =   Centroid Deltas}

 {\em Summary:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and
where you want to be  (which we will call $C_j$).

 {\em Details:} Method1 clusters project data by reflecting   on the independent variables, then
  reporting the delta between the cluster centroids. 
  After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
continuous/discrete feature. After that, it
finds the closest centroid $C_j$ that has a better
performance score. For defect data, ``better'' means fewer defective examples while for the config data,
``better'' means lower median runtimes for the examples in that cluster.
Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

\subsection{Method2 = \\\CD+FS =  Method1  + Feature Selection }

{\em Summary:} Method2 works like Method1, but this time, it only mentions the most
informative features. Hence, Method2's plans are simpler.

{\em Details:} In most SE applications, not all features contribute equally to a problem.  Hence, one problem
with the above method is that its plans comment on all features, even if those values are due
to noise or other effects not connected to better performance.

This is a standard issue in data mining, with a standard solution.
A common result is that the signal in a table of data is mostly contained in a handful of features~\cite{hall03,kohavi97}.
Papakroni~\cite{papa13} has tested for this effect in the Jureczko data sets.
Papakroni found no loss of   efficacy in defect prediction after
sorting all features by their information content,
then making predictions using (a)~all  features or (b)~just using   33\% most informative features.

Based on the above, it might be possible to simplify Method1's plans by pruning back the features used to make
the plans. Following on from Papakroni, our Method2 returns plans
containing just the top $\beta=33\%$ most informative features. Here, ``informative'' means
that the values of a feature are good for selecting a small set of clusters (ideally,
just one).
This can be estimated using the Fayyad-Iranni INFOGAIN algorithm~\cite{FayIra93Multi}
as follows.

Firstly, 
run  WHERE to generate   clusters/ Next,
discretize numeric features using the Fayyad-Iranni discretizer
(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
and 
let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
Cull the features as per Papakroni; i.e. just use the $\beta=33\%$ most informative features
where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).

After that, Method2 works line Method1 but now the   plans  only
mention the $\beta=33\%$ most informative features.
\subsection{Method3 = BIC = Method2 + Best-in-cluster}\label{sect:BIC}

{\em Summary:} Method3 is like Method2, but it uses more knowledge about the training data.

{\em Details:}
One problem with Method2 is that it makes little use of the detailed knowledge
of the examples that are squashed together to form the centroids. Accordingly,
our Method3 works like Method2 but some important differences.

In Method3, clusters are summarized in {\em two} examples: (1)~the centroid found in Method1 and
(2)~the best-in-cluster example; i.e. the  example in that cluster
with the best performance score.
Next, as before, each centroid $C_i$ is annotated with the nearest centroid $C_j$
with a better performance score.
Howver, after that,  when generating a plan for a test instance, Method3:
\bi
\item
Find the test cases's nearest centroid $C_i$;
\item
Finds the associated better centroid $C_j$ and its associated best-in-cluster example.
\item
Computes a plan to change the test case  from the delta between the test case
and the best in that $C_j$ cluster.
\ei
Note that, as with Method2, this method's plans only mentions the  $\beta=33\%$ most informative features. 
%\input{tex/pseudo-code-1.tex}

\input{tex/dtree.tex}

\subsection{Method4 = XTREES = Deltas in Decision Tree Branches}

{\em Summary:} Method4 builds a decision tree,  then generates
plans from the difference between two branches:
(1)~the branch to where you are and (2)~the branch to where you want to be.

{\em Details:} One potential problem with Methods 1,2 and 3 is the {\em unsupervised} nature of
the algorithm used for the initial clustering. Unsupervised methods like WHERE
execute without knowledge of the target class.
{\em Supervised} methods, on the other hand, are allowed to reflect on more knowledge including
the values in the target class. In theory, this extra knowledge can lead to better data mining.
Accordingly, Method4 uses a supervised method to generate its plans.

In summary, XTREES uses the decision tree algorithm of \fig{where}.D to divide the data.
Decision trees have branches down to leaves. 
XTREES builds plans by asking three questions:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee
The last question can be answered by finding the deltas in branches of the decision tree that lead to {\em desired} from {\em current}. For full algorithmic details, refer to \fig{xtrees_bare}.

\input{tex/results_models.tex}

\section{Experiments}

\begin{figure}[!t]
\small 
\[
\begin{array}{r}
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; HOW)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{actual}& =\mathrm{Defects \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
           \mathrm{\bf if\;}\mathit{actual} & >  \mathit{0}\\
           \mathrm{\bf then} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{actual},\mathit{after}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
\end{figure}

\subsection{Setup}

In the experiments shown below,  we use the planners to comment on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be manually adjusted by programmers or (for large scale changes) automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}.

\fig{work} shows our proposal to verify if the recommendations made by our planners actually work. Accordingly, we divide the
project data  into two disjoint sets {\em train} and {\em test}
(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
Next, from the train set, we build both a {\em planner} as well
as a {\em  predictor}. Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper, our {\em planner} will be from \ref{clust_contrast} and our {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} and Random Forest Regressor taken from the SciKit Learn toolkit~\cite{Pedregosa2012}.

As for the {\em test} data, this is passed to the { predictor}
to measure the baseline prediction accuracy, false alarm, and recall.
If our { predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data $test'$;
then (3)~return data on the {\em before, after} predictions.

\subsection{Data}

\input{tex/results_jur.tex}
\input{tex/results_conf.tex}

\begin{figure}[t]
\begin{minipage}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{_figs/Deltas-Lucene.eps}
\label{fig:delta_ivy}
\end{minipage}
\begin{minipage}{0.5\textwidth}
\includegraphics[width=0.9\linewidth]{_figs/Deltas-BDBJ.eps}
\label{fig:delta_apache}
\end{minipage}
\caption{Performance Comparison showing the number of times a certain feature requires to be changed. X-axis lists the various features that make up a data set. The y-axis contains the frequency of change in \%.}
\end{figure}

The clear advantage of our current work is that it can be readily and automatically applied to domains with historical data. However, the obvious drawback with this approach is that our assessment of the plans is only as good as the predictor. To mitigate for this {\em problem of poor predictors}, it is wise to demand certain standards for {\em satisfactory} thresholds.

\subsection{Assessment}
\fig{work} showed how we build and test our planners. The dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data. Additionally, for both the data sets, we designed a setup that repeated the trials 40 times (the value 40 was chosen to satisfy the minimum sample size requirements of the central limit theorem):
\bi
\item For Siegmund data, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data.
\item For the Jureczko data, we used the training and testing sets of \fig{j}. 
\ei
The 40 repeats defined above were repeated for each planner. A Scott-Knott procedure was used to test for statistical differences in the result. The Scott-Knott procedure was recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{mittas13}.

This method
sorts a list of $l$ treatments with $ls$ measurements by their median
score. It then
splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 $E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.
Scott-Knott  applies an  hypothesis test $H$ to check
if $m,n$ are significantly different. If so, Scott-Knott  recurses on the splits.

The hypothesis test $H$ used here is a conjunction of the A12 effect size test and a non-parametric bootstrap sampling. In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$). For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}. For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}. These researchers warn that even if an hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small. Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.

\subsection{Results}
\input{tex/deltas_avg.tex}
The results for each data, for different planners, are shown in \fig{jur} and \fig{conf}. All
\subsection{Discussion}	whye where?
Recent results suggest that, when clustering on independent variables of software project data,
	we are free to select from a wide range of 
	clustering methods.  Ganesan~\cite{div14} explored 
	different clustering methods for SE data using   effort and defect data from
	the PROMISE repository~\cite{promiserepo}.
	That study explored
	K-Means, mini-batch-K-Means, DBscan, EM, Ward, and the WHERE algorithm discussed
	below.
	Clusters were assessed via the performance of prediction 
	learned from the clusters by Random Forests (for defect prediction)
	and M5prime or linear regression (for effect estimation).  Ganesan found
	that the details of the clustering method were less important than ensuring that  a large number of clusters were generated.
	Accordingly, for clustering, we select a clustering method that generates a large
	number of small cluster
	($\sqrt{N}$ clusters
	from $N$ examples) that runs fasts, which has shown promise in prior work~\cite{Menzies2013}, and which ignores spurious dimensions (the last item is important since, in our experience, much SE data is ``noisey''; i.e. contains signals not associated with the target variable.

\section{Threats to Validity}
\section{Related Work}
dashboards

strucftureal learners (bayes)

\section{Conclusion}

works for more than defects and tunrimes

This main contribution of this paper was to  define an evaluation strategy for this kind of reasoning, and to show that (using that strategy) we
can (a)~select useful planning methods while (b)~culling clearly
inferior planning methods. 
We apply this strategy to XTREES and three alternate planning methods to find:
\bi
\item Two methods that are clearly inferior;
\item One method that proposes complex changes, which work as well as XTREES;
\item And XTREES, that proposes very simple changes.
\ei
Of course, there are many more methods for generating plans and
no   one paper can survey them all.
But the goal of this paper is not to claim that (e.g.) XTREES is some absolute optimal algorithm. Rather, it is
to offer a baseline result (with XTREES) and an  evaluation strategy that  can assess  if alternate methods are better than XTREES.
Our hope is that other   researchers apply this strategy  to repeat and/or improve
and/or refute our results. To that end, all our materials and data sets are available on-line, see
github.com/ai-se/XTREES.

\section*{Acknowledgements}
The work has partially funded by a National Science Foundation CISE CCF award \#1506586.
\bibliographystyle{plain}

\bibliography{References}
\end{document}

  Next, build one centroid for each cluster (using the median and mode value for continuous
and discrete values, respectively).
After that, write a spreadsheet with one column per centroid;
\bi
\item
Sort the columns such that any current projects client fall into the left-most
columns. That is, make the left-hand-side of the sheet  ``current usual practice''
and the right-hand side  ``alternatives to current usual practice''.
\item
Sort the rows of that spreadsheet such that the rows with
maximum  variability appear at the top 
\ei
Now show the sheet to the business user, encouraging them to make recommendations by
\bi
\item
Comparing the left and right-sided columns;
\item
Using the features mentioned in the top-most rows (where changes selects for the most different centroids).
\ei

In theory, One  advantage of this method is that it focuses the attention of the business users
away from rows that do not select for different centroids and towards differences between
current usual practice and everything else.  
In the summer of 2011 and 2012, one of us (Menzies) spent two months
	working on-site at Microsoft Redmond,
	observing data mining analysts.  In that study, he took special
	note about how Microsoft's data scientists
	discussed the results of their data mining sessions with  business users.

	One surprising observation was how  
	little time was spent by business users 
	inspecting  of the output of standard data miners. Prior to that visit,
	we had the mistaken impression that   decision trees,
	clustering algorithms, etc were useful ``off-the-shelf''; i.e    business
	users would inspect and understand the output of those tools.

Standard data mining tools are not necessarily the best tool for supporting that dialogue.
Menzies found that he had to do considerable work pre-processing data mining output
prior to the weekly briefing meetings for the business users. Initially,
that pre-processing was just clustering. This evolved into feature selection and finally
the a case-based reasoning tool called HOW.

	as compared to another process, which we call {\em peeking}.
	In {\em peeking}, analysts and users spend much time
	inspecting and discussing small samples of either raw or exemplary or synthesized project data.  Further, very little of those discussions were  focused on classification
	(the addition of a labels to some unlabelled data). Rather, much time
	was spent in those meetings discussing {\em what to do next}; i.e. trying
	to determine what could be altered to better improve some business outcome.

	That   Microsoft  study found two common ``peeking'' methods.
	In {\em data engagement meetings},
	users debated the implications of data
	displayed on a screen. In this way, users
	engaged with the data and with each other by
	monitoring each others' queries and check each others'
	conclusions.

	Another data analysis pattern observed
	at Microsoft was  {\em cluster + contrast} in which
	data is  reduced to a few
	clusters. Users are then just shown the delta between those
	clusters. While contrasting, if feature values are
	the same in both clusters, then these were pruned from
	the reports to the user. In this way, very large
	data sets can be shown on one PowerPoint
	slide. Note that {\em cluster+contrast} is a tool that can be usefully employed within
	{\em data engagement meetings}.

	Cluster+contrast and engagement
	meetings are common practices at Microsoft. Yet  these methods had never been rigorously studied or certified.
	For both those reasons,
	we reflected over those tools to discover and analyze their
	underlying process. The result was HOW~\cite{howase}: a tool
	that combines (a)~feature selection; (b)~centroid generation from   clusters;
	(c)~contrast methods between centroids.
	While method (a) is widely used (e.g.~\cite{Menzies2010}),
	to the best of our knowledge, this combination of (abc) has not been thoroughly explored before.

This paper assesses two methods  for ``peeking'': a model-based method called DTREE and an instance-based method   called HOW. 
DTREE were first developed by Lekkalapudi and Menzies to  explaining results from multi-objective optimizers~\cite{nva14}. This paper is the first
to apply DTREE to defect prediction and contrast set learning. Also, that prior work evaluated
DTREE against multi-objective optimizers and not   instance-based methods.

This paper uses three criteria to assess the value of HOW and DTREE for
learning actionable analytics:
\be
\item
A planning system needs to be {\em effective}; i.e. if its recommendations
are applied then some statistically significant change should be observed.
To predict the number of defects in a data set before and after applying our changes,
we build a    prediction system (built by data mining; specifically: Random Forest). Note that this
predictor was built from some hold-out data (i.e. from  different data than that used
to build the predictor).
\item
Any  conclusion made to a business user must be understandable;
it must generate {\em succinct} changes.  
\item
Recommendations should be {\em stable}; i.e. they shouldn't widely vary due to minor changes in the data. Hence, in our experiments, we will add a little randomness to our analysis then report results across 20 repeated runs. 
\ee
We will find that
 {\em effectiveness} of DTREE and HOW are similar, but   DTREE wins on {\em succinctness} and {\em stability}.

\section{Cluster and Contrast}
\label{clust_contrast}
A recurring data analysis pattern in this paper is $cluster+contrast$. The data is distilled into a few clusters using a clustering scheme. Then lessons are inferred by studying the differences between these clusters. These lessons are used to generate rules that can be applied in any context.

\subsection{Clustering}
There are a wide variety of clustering methods to choose from. A study by Ganesan~\cite{div14} explored different clustering methods for software engineering data using the effort and defect data from the PROMISE repository~\cite{promise}. In that study methods such as WHERE, K-Means, mini-batch-K-Means, DBScan, EM, and Ward were investigated. The results of the study showed that the size and number of clusters is more important that the specifics of the techniques used.

For this purposes of this work, we have chosen WHERE, a clustering scheme which is capable of generating at least $\sqrt{N}$ clusters given $N$ instances. In addition to this, WHERE has been shown to run fast while ignoring spurious dimensions~\cite{menzies2013}. This is particularly useful, for much of the SE data is noisy, they contain a information not associated with the target variable.

\subsection{Finding Contrasts}
All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability. It works as follows:
\bi
\item Find   two   distance samples from the data, say  $X,Y$. This can be done by  picking any case $W$ at random, then setting $X$ to its most distant case, then setting $Y$ to the case most distant from $X$ (this requires only $O(2N)$ comparisons of $N$ cases).
\item Project each case $Z$ onto a {\tt Slope} that  runs between $X,Y$ using the cosine rule. 
\item Split the data at the median $X$ value of all cases and recurse on each half  (stopping when one half has less  than $\sqrt{N}$ of the original population).
\ei

Clustering is followed by generating \textit{contrast sets}. These contrast sets represent recommendations on what could be altered to better improve an outcome. In this work we have explored several algorithms as possible tools to identify contrast between clusters. These fall into two broad categories:
\begin{enumerate}
\item Case Based Reasoning techniques (Nearest Neighbors and a gradient base planner called HOW)
\item Decision Trees.
\end{enumerate}

These techniques are discussed below. It is worth noting that the following techniques are organized in such a way that each one seeks to address certain shortcomings in the ones that precede it.

\subsection{Case Based Reasoning}

Case-based reasoning seeks to find solutions to problems by emulating human recollection and adaptation from past experiences. It has found extensive usage in Artificial Intelligence because it offers several advantages. However, one of the most important benefits that CBR has to offer is that it works on a ``case-by-case'' basis. Therefore it's advise is tailored to be specific to the particular case being considered. Several paper in SE have applied this technique, most of all for effort estimation ~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}.

A classic example of a CBR is K Nearest Neighbor. Our version of nearest neighbor for planning is rather straight forward. It has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.

\subsection{HOW}

HOW is very different compared to conventional CBR planners in that it explores the gradient between pairs of nearby clusters instead of studying the clusters themselves. HOW works by clustering the data during training using WHERE and then drawing \texttt{slopes} between the centroids of pairs of nearby clusters. Assuming the cluster pairs are labeled X and Y, with X having slightly better performance score than Y, the \texttt{slope} between X and Y acts as an indicator pointing to a direction in which to displace the data; i.e. away from Y and towards X.

While testing, HOW finds the nearest slope to every test case. The slope provides the exact magnitude and direction of displacements. Contrast sets are derived from these displacements. HOW offers a distinct advantage over the other CBR planners by limiting the displacements to very small regions (the displacements are never more than the separation between two clusters).

Although HOW manages to generate plans by localizing displacements to regions small enough to produce a statistically significant improvement, it needs to be noted that it fails to provide succinct summaries. Lack of succinctness makes it difficult draw generalizable conclusions about the test data. This is a trend commonly observed in most CBR systems. They tend to reason directly from a loaded training data instead of first summarizing the data into a model.

% However, not all domains come with reliable models. For instance, a model that encompasses all the intricate issues that may lead to defects in software would be very large and indeed rather complex. In addition to this, finding empirical data to validate such models can be hard to come by and also time consuming ~\cite{me09i,me09j}.

\subsection{DECISION TREES}

Trustworthy domain models are a popular alternatives to CBR planners. In our previous work, we have used executing source code as a "model" to check if our mutations to test suites minimize the test suite size while maximizing the number of statements covered~\cite{me09m,andrews07,andrews10}. However, we do not always have access to ready-to-use models. Further, finding empirical data to validate existing models can be hard.

Taking into account the wealth of data that is available to us, this issue can fortunately be circumvented by constructing a categorical model based on decision trees. Unlike the previous planners which use where, here we construct a decision tree using information gain as a metric to find ideal splits in the data. This allows us to identify the most informative attributes to select. A decision tree (hereafter referred to as DTREE) built in this fashion emulates a model built from the training data.

Using DTREE, the test cases can be categorized into one of the branches of the tree. Now, to generate the contrast sets, we determine (1) What \textit{current} branch does a test case fall in?; (2) What \textit{desired} branch would the test case have to move to?; (3) What are the deltas between \textit{current} and \textit{desired}? The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}. For full algorithmic details, refer to \fig{xtrees_bare}.

A tree structure such as DTREE is characterized by attributes such as size and depth. It is worth noting that these attributes have a profound impact on its performance. Our initial motivation for using DTREE was that they could serve as a medium for experts to reason about a data. The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity. We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare}).

DTREE offers a great number of benefits compared to the other methods that were discussed above. Firstly, it offers a visual medium for experts to identify and explore solutions spaces that are local to the problem. Secondly, its solutions are much more stable than other cluster specific and instance specific learners. This stability can be attributed to the consistency and general reproducibility of the tree structure. Thirdly, and perhaps most importantly, the tree summarizes the training data succinctly, making it easier for a business user to examine the solution sets.

\lstnewenvironment{code}[1]{
\lstset{
mathescape,
numbers=left,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=1em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize,
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, FSel, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}