### 2015-08-24 tex/algos.tex

	\begin{figure} 
	\begin{shaded}
  ~\hrule~

 {\bf Figure 7.A: Measuring Variability}

	For  continuous and discrete values,
the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{x_i - \bar{x})^2}{n-1}\right)}$}
where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
for  $n$ discrete values  at frequency
$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.

  ~\hrule~

	{\bf Figure 7.B: Measuring distance}

	We use
		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  w_i\Delta(X_i,Y_i)^2}$. Here, $w_i$ is a weight term for each feature
		 (usually set to 1).
			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
			Otherwise, replace any  missing items with values that maximizes the following.
			For numerics $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
			returns 
			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
			same,different (respectively).

 ~\hrule~

	{\bf Figure 7.C: Top-down Clustering with WHERE}

				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$ using 
		Using this measure, WHERE runs as follows:
		\begin{enumerate}[leftmargin=3mm]
			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(which requires only $O(2N)$ comparisons).
			\item Project each case $Z$
			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
			\item Split the data at the median $x$ value of all cases.
			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse   from step1.
		\end{enumerate}
 In terms of related work,
	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
	  at each recursive level to find the PCA principle component. Our method, on the other hand,
	  performs the same task with only $O(2N)$ distance calculations 	using the 
	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.

	   ~\hrule~

	{\bf Figure 7.D: Top-down division with Decision Trees}

	Find a split in the values of  independent features that most reduces the variability
  of the dependent feature (measured using \fig{where}.A). Divide data on that  feature's split and 
   for all splits bigger than $\alpha=\sqrt{N}$,  recurses on each split.

	 ~\hrule~

	 	{\bf Figure 7.E: Finding the most informative rows}

Discretize all numeric features using the Fayyad-Iranni discretizer~\cite{FayIra93Multi}
(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
and 
let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
Cull the the features as per Papakroni~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).

~\hrule~

	  \end{shaded}
		\caption{Some algorithms used in this paper.}\label{fig:where}
		\end{figure}