### 2015-08-24 tex/results_models.tex

\begin{figure}[htbp]

~\hrule~

%\noindent\begin{minipage}{0.5\textwidth} 
{\small 
\begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
  1 &        XTREE &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
\hline  2 &      CD+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
  2 &          BIC &    1.04  &  0.01 & \quart{42}{0}{42}{20} \\
  2 &      CD &    1.11  &  1.24 & \quart{18}{80}{46}{20} \\
\hline 
\end{tabular}}

 ~\hrule~
\caption{Results: Methods 1,2,3,4 applied to some ground-truth data.
Values collected from   20 repeated runs of each method with different random seeds.
}\label{fig:conf}
\end{figure}