### 2015-08-25 tex/results_models.tex

\begin{figure}[!b]

~\hrule~

%\noindent\begin{minipage}{0.5\textwidth} 
{\small 
\begin{tabular}{l@{~~~}l@{~~~}r@{~~~}r@{~~~}c}
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
  1 &        XTREE &    0.42  &  0.17 & \quart{0}{10}{0}{20} \\
\hline  2 &      CD+FS &    1.03  &  0.69 & \quart{30}{46}{42}{20} \\
  2 &          BIC &    1.04  &  0.01 & \quart{42}{0}{42}{20} \\
  2 &      CD &    1.11  &  1.24 & \quart{18}{80}{46}{20} \\
\hline 
\end{tabular}}

 ~\hrule~
\caption{Results: Methods 1,2,3,4 applied to some ground-truth data.
Values collected from   40 repeated runs of each method with different random seeds.
Results show ratios of (1) sum of development effort (in the changes examples) to (2)
the of the development effort in the original test data.
}\label{fig:coc}
\end{figure}