### 2015-08-27 tex/results_models.tex

\begin{figure}[!t]
%\noindent\begin{minipage}{0.5\textwidth} 
{\small 
\begin{tabular}{llrrc}
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{50th} & \textbf{25-75th} & \\\hline
  1 &          HOW &    -3.3  &  2.2 & \quart{26}{2}{27}{32} \\
  1 &      baseln0 &    -2.9  &  11.7 & \quart{17}{14}{28}{32} \\
  1 &      baseln1 &    -2.75  &  65.5 & \quart{0}{78}{28}{32} \\
\hline  2 &        dtree &    51.5  &  14.84 & \quart{82}{17}{93}{32} \\
\end{tabular}}
\caption{Methods 1,2,3,4 applied to some ground-truth data (in this case, the POM3 model).
Values collected from  40 repeated runs of each method with different random seeds.
Results show the efficacy of XTREE in reducing the total overall cost in the original test data, when other planners fail to do so.
}\label{fig:coc}
\end{figure}